---
title: "dplyr_Lab_Assignment"
author: "Po Wei Tsao"
date: "9/7/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
getwd()

```


For this assignment you can use any or all of the datasets I 
provided or choose three of your own!  I would like you to compose 
two questions for each dataset that requires at a minimum three "levels" 
of pseudocode and use dplyr functions to solve the statements. Also create 
functions for each of your  questions. Do not repeat function sequences or questions.  Information or the actual data for the datasets can be found through the links below. In summary for each dataset produce five parts: 1)Question, 2)psuedocode, 3)single line
using dplyr functions, 4) pipe the individual lines together and 
5) create a function. 

Also you can work with your groups for coding tips/advise or work through similar programming issues but everyone must have their own set of questions and results. 


# beaches data is in the working directory 

# nf2008_fg: http://users.stat.ufl.edu/~winner/data/nfl2008_fga.csv

# red_wine_quality: https://data.world/uci/wine-quality


# Lets walk through a example. Below is a dataset on students admitted to a graduate program in Data Science (actually don't remember where this data came from). 

## Are there differences in the admitted number of students on average by GPA and school rank?

  >A easy way to handle this is to identify function calls starting at the end of the question and working backwards. 

## Pseudocode 
  - Group by school rank
  - Calculate GPA Average
  - Count total admitted 
  - Calculate % Admitted
  - Arrange in descending order by Mean GPA


Loading and preparing the data 
```{r}
admit_2 <- read_csv("DS-3001/data/LogReg.csv")

#creating a quick list to use as a index
names <- c("admit","rank")

names
#changing those two variables to factors
admit_2[,names] <- lapply(admit_2[,names],factor)

str(admit_2)

```

Single line version 
```{r}
#Group
x1 <- group_by(admit_2,rank)
#Calculate additional variables
x2 <- summarize(x1,count=n(),ave=mean(as.integer(admit)-1),mean(gpa))
#Arrange by GPA
x3 <- arrange(x2,desc(`rank`))
print(x3)

```

## Piping it all together
```{r}
#creating a dplyr version
(admit_1 <- group_by(admit_2, rank)%>%
  summarize(count=n(),ave=mean(as.integer(admit)-1),mean(gpa))%>%
   arrange(desc(`mean(gpa)`)))
```


## Now just turn it into a function, so if you needed to repeat this data request on future datasets it would be simple. 
```{r}
ave_admit <- function(x){
  xx <- group_by(x, rank)%>%
  summarize(count=n(),ave=mean(as.integer(admit)-1),mean(gpa))%>%
  arrange(desc(`mean(gpa)`))  
  print(xx)
}

ave_admit(admit_2)

```


##Dataset 1, question 1
```{r}
# question: Are there differences in rainfall based on seasons?

#pseudocode: 
#1. Group by seasons
#2. calcualate average rainfall
#3. Order by average rainfall

#line by line
beaches <- read_csv("ds3001-personal/data/beaches.csv")
beaches = na.omit(beaches)
seasons <- group_by(beaches, season_name)
avg = summarize(seasons,ave_rainfall=mean(rainfall))
avg_rainfall_by_season = arrange(avg, desc(ave_rainfall))
print(avg_rainfall_by_season)
```

```{r}
#function
beaches <- read_csv("ds3001-personal/data/beaches.csv")
beaches = na.omit(beaches)

ave_rainfall <- function(beaches){
avg_rainfall_by_season <- group_by(beaches, season_name)%>%
  summarize(ave_rainfall=mean(rainfall))%>%
   arrange(desc(ave_rainfall))
  print(avg_rainfall_by_season)
}

ave_rainfall(beaches)
```

## Dataset 1, question 2

```{r}
#question: Are there any trends between temperature and season?

#pseudocode:
#1. Group by seasons
#2. calcualate average temperature
#3. Order by average temperature

#line by line
beaches <- read_csv("ds3001-personal/data/beaches.csv")
beaches = na.omit(beaches)

season_names = group_by(beaches, season_name)
summary = summarize(season_names, ave_temp=mean(temperature))
avg_temp_by_season = arrange(summary, desc(ave_temp))
print(avg_temp_by_season)

```

```{r}
#function
beaches <- read_csv("ds3001-personal/data/beaches.csv")
beaches = na.omit(beaches)

ave_temp <- function(beaches){
avg_temp_by_season <- group_by(beaches, season_name)%>%
  summarize(ave_temp=mean(temperature))%>%
   arrange(desc(ave_temp))
  print(avg_temp_by_season)
}

ave_temp(beaches)
```


##Dataset 2, question 1
```{r}
# question: how many times were drivers of each race stopped?

#pseudocode: 
#1. Group by Driver_Race
#2. calcualate total stops for each race
#3. Order by total stops

#line by line
stops <- read_csv("ds3001-personal/data/officer_Traffic_Stops.csv")
stops = na.omit(stops)

race = group_by(stops, Driver_Race)
summary = summarize(race, stop_count=n())
stop_count = arrange(summary, desc(stop_count))
print(stop_count)

```

```{r}
#function
stops <- read_csv("ds3001-personal/data/officer_Traffic_Stops.csv")
stops = na.omit(stops)
stop_count = function(stops){
(result <- group_by(stops, Driver_Race)%>%
  summarize(stop_count=n())%>%
   arrange(desc(stop_count)))
  print(result)
}

stop_count(stops)
```

##Dataset 2, question 2
```{r}
# question: how many times was a search conducted for drivers of each race?

#pseudocode: 
#1. Get all rows where a search was conducted
#2  Group by Driver_Race
#2. calcualate total searches for each race
#3. Order by total searches

#line by line
stops <- read_csv("ds3001-personal/data/officer_Traffic_Stops.csv")
stops = na.omit(stops)

search_conducted = subset(stops, Was_a_Search_Conducted == "Yes")
race = group_by(search_conducted, Driver_Race)
summary = summarize(race, search_count=n())
search_count = arrange(summary, desc(search_count))
print(search_count)

```

```{r}

#function 
stops <- read_csv("ds3001-personal/data/officer_Traffic_Stops.csv")
stops = na.omit(stops)
get_total_searches = function(stops){
  search_count <- subset(stops, Was_a_Search_Conducted == "Yes")%>%
    group_by(Driver_Race)%>%
    summarize(search_count=n())%>%
    arrange(desc(search_count))
  print(search_count)
}
get_total_searches(stops)

```

##Dataset 3, question 1

```{r}
#question: Are there differences in average account balance between people with different jobs

#pseudocode
#1. Group by job
#2. Calculate average account balance
#3. Arrange by descending average account balance

#line by line
bank <- read_csv("ds3001-personal/data/bank.csv")
bank = na.omit(bank)
jobs = group_by(bank, job)
summary = summarize(jobs, avg_acc_balance=mean(balance))
avg_acc_balance_by_job = arrange(summary, desc(avg_acc_balance))
print(avg_acc_balance_by_job)

```

```{r}
#function
bank <- read_csv("ds3001-personal/data/bank.csv")
bank = na.omit(bank)
get_avg_balance_by_job = function(bank){
  (avg_acc_balance_by_job <- group_by(bank, job)%>%
  summarize(avg_acc_balance=mean(balance))%>%
   arrange(desc(avg_acc_balance)))
  print(avg_acc_balance_by_job)
}
get_avg_balance_by_job(bank)

```


##Dataset 3, question 2

```{r}
#question: Are there differences in average account balance between people with different ages

#pseudocode
#1. Group by age
#2. Calculate average account balance
#3. Arrange by descending average account balance

#line_by_line
bank <- read_csv("ds3001-personal/data/bank.csv")
bank = na.omit(bank)
ages = group_by(bank, age=cut(age, breaks=seq(0, 100, by=10)))
summary = summarize(ages, avg_acc_balance=mean(balance))
avg_acc_balance_by_age = arrange(summary, desc(avg_acc_balance))
print(avg_acc_balance_by_age)

```

```{r}
#function
bank <- read_csv("ds3001-personal/data/bank.csv")
bank = na.omit(bank)
get_avg_balance_by_age = function(bank){
(avg_acc_balance_by_age <- group_by(bank, age=cut(age, breaks=seq(0, 100, by=10)))%>%
  summarize(avg_acc_balance=mean(balance))%>%
   arrange(desc(avg_acc_balance)))
  print(avg_acc_balance_by_age)
}
get_avg_balance_by_age(bank)

```
