---
title: "dplyr_Lab_Assignment"
author: "Po Wei Tsao"
date: "9/7/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


For this assignment you can use any or all of the datasets I 
provided or choose three of your own!  I would like you to compose 
two questions for each dataset that requires at a minimum three "levels" 
of pseudocode and use dplyr functions to solve the statements. Also create 
functions for each of your  questions.  Information or the actual 
data for the datasets can be found through the links below. In summary for
each dataset produce five parts: 1)Question, 2)psuedocode, 3)single line
using dplyr functions, 4) pipe the individual lines together and 
5) create a function. 


##Dataset 1, question 1
```{r}
# question: Are there differences in rainfall based on seasons?

#pseudocode: 
#1. Group by seasons
#2. calcualate average rainfall
#3. Order by average rainfall

#line by line
beaches <- read_csv("ds3001-personal/data/beaches.csv")
beaches = na.omit(beaches)
seasons <- group_by(beaches, season_name)
avg = summarize(seasons,ave_rainfall=mean(rainfall))
avg_rainfall_by_season = arrange(avg, desc(ave_rainfall))
print(avg_rainfall_by_season)
```

```{r}
#function
beaches <- read_csv("ds3001-personal/data/beaches.csv")
beaches = na.omit(beaches)

ave_rainfall <- function(beaches){
avg_rainfall_by_season <- group_by(beaches, season_name)%>%
  summarize(ave_rainfall=mean(rainfall))%>%
   arrange(desc(ave_rainfall))
  print(avg_rainfall_by_season)
}

ave_rainfall(beaches)
```

## Dataset 1, question 2

```{r}
#question: Are there any trends between temperature and season?

#pseudocode:
#1. Group by seasons
#2. calcualate average temperature
#3. Order by average temperature

#line by line
beaches <- read_csv("ds3001-personal/data/beaches.csv")
beaches = na.omit(beaches)

season_names = group_by(beaches, season_name)
summary = summarize(season_names, ave_temp=mean(temperature))
avg_temp_by_season = arrange(summary, desc(ave_temp))
print(avg_temp_by_season)

```

```{r}
#function
beaches <- read_csv("ds3001-personal/data/beaches.csv")
beaches = na.omit(beaches)

ave_temp <- function(beaches){
avg_temp_by_season <- group_by(beaches, season_name)%>%
  summarize(ave_temp=mean(temperature))%>%
   arrange(desc(ave_temp))
  print(avg_temp_by_season)
}

ave_temp(beaches)
```


##Dataset 2, question 1
```{r}
# question: how many times were drivers of each race stopped?

#pseudocode: 
#1. Group by Driver_Race
#2. calcualate total stops for each race
#3. Order by total stops

#line by line
stops <- read_csv("ds3001-personal/data/officer_Traffic_Stops.csv")
stops = na.omit(stops)

race = group_by(stops, Driver_Race)
summary = summarize(race, stop_count=n())
stop_count = arrange(summary, desc(stop_count))
print(stop_count)

```

```{r}
#function
stops <- read_csv("ds3001-personal/data/officer_Traffic_Stops.csv")
stops = na.omit(stops)
stop_count = function(stops){
(result <- group_by(stops, Driver_Race)%>%
  summarize(stop_count=n())%>%
   arrange(desc(stop_count)))
  print(result)
}

stop_count(stops)
```

##Dataset 2, question 2
```{r}
# question: how many times was a search conducted for drivers of each race?

#pseudocode: 
#1. Get all rows where a search was conducted
#2  Group by Driver_Race
#2. calcualate total searches for each race
#3. Order by total searches

#line by line
stops <- read_csv("ds3001-personal/data/officer_Traffic_Stops.csv")
stops = na.omit(stops)

search_conducted = subset(stops, Was_a_Search_Conducted == "Yes")
race = group_by(search_conducted, Driver_Race)
summary = summarize(race, search_count=n())
search_count = arrange(summary, desc(search_count))
print(search_count)

```

```{r}

#function 
stops <- read_csv("ds3001-personal/data/officer_Traffic_Stops.csv")
stops = na.omit(stops)
get_total_searches = function(stops){
  search_count <- subset(stops, Was_a_Search_Conducted == "Yes")%>%
    group_by(Driver_Race)%>%
    summarize(search_count=n())%>%
    arrange(desc(search_count))
  print(search_count)
}
get_total_searches(stops)

```

##Dataset 3, question 1

```{r}
#question: Are there differences in average account balance between people with different jobs

#pseudocode
#1. Group by job
#2. Calculate average account balance
#3. Arrange by descending average account balance

#line by line
bank <- read_csv("ds3001-personal/data/bank.csv")
bank = na.omit(bank)
jobs = group_by(bank, job)
summary = summarize(jobs, avg_acc_balance=mean(balance))
avg_acc_balance_by_job = arrange(summary, desc(avg_acc_balance))
print(avg_acc_balance_by_job)

```

```{r}
#function
bank <- read_csv("ds3001-personal/data/bank.csv")
bank = na.omit(bank)
get_avg_balance_by_job = function(bank){
  (avg_acc_balance_by_job <- group_by(bank, job)%>%
  summarize(avg_acc_balance=mean(balance))%>%
   arrange(desc(avg_acc_balance)))
  print(avg_acc_balance_by_job)
}
get_avg_balance_by_job(bank)

```


##Dataset 3, question 2

```{r}
#question: Are there differences in average account balance between people with different ages

#pseudocode
#1. Group by age
#2. Calculate average account balance
#3. Arrange by descending average account balance

#line_by_line
bank <- read_csv("ds3001-personal/data/bank.csv")
bank = na.omit(bank)
ages = group_by(bank, age=cut(age, breaks=seq(0, 100, by=10)))
summary = summarize(ages, avg_acc_balance=mean(balance))
avg_acc_balance_by_age = arrange(summary, desc(avg_acc_balance))
print(avg_acc_balance_by_age)

```

```{r}
#function
bank <- read_csv("ds3001-personal/data/bank.csv")
bank = na.omit(bank)
get_avg_balance_by_age = function(bank){
(avg_acc_balance_by_age <- group_by(bank, age=cut(age, breaks=seq(0, 100, by=10)))%>%
  summarize(avg_acc_balance=mean(balance))%>%
   arrange(desc(avg_acc_balance)))
  print(avg_acc_balance_by_age)
}
get_avg_balance_by_age(bank)

```
